### 2019-06-18 text/properties.tex
Revise Section 4.

% !TEX root = ../main.tex

\section{Properties of \LMD}

In this section, we show basic properties of \LMD: preservation, strong normalization, confluence for full reduction, and progress for staged reduction.

% Substitution Lemma

Substitution Lemma in \LMD{} is little more complicated than usual
because there are eight judgment forms and two kinds of substitution.
Term Substitution Lemma states that term substitution $[z \mapsto M]$
preserves derivability of judgments and Stage Substitution Lemma
states similarly for stage substitution $[\alpha\mapsto A]$.

We let $\mathcal{J}$ stand for judgments $K \iskind @A$, $\tau::K@A$,
$M:\tau@A$, $K \E J @ A$, $\tau \E \sigma @ A$, and
$M \E N : \tau @ A$.  Substitutions $\mathcal{J}[z \mapsto M]$ and
$\mathcal{J}[\alpha \mapsto A]$ are defined in a straightforward
manner.  Using these notations, the two substitution lemmas are stated as follows:
\AI{Did we define $\Delta [x \mapsto N]$?}

\begin{lemma}[Term Substitution]
  If $\G, z:\xi@B, \D \V \mathcal{J}$ and $\G\V P:\xi @B$, then $\G, \D[z \mapsto P] \V \mathcal{J}[z \mapsto P]$.  Similarly, if $\V \G, z:\xi@B, \D$ and
  $\G\V P:\xi @B$, then $\V \G, \D[z \mapsto P]$.
\end{lemma}

\red{TODO: Fix the proof.}
\red{$\G\V J@A$における代入を定義する必要があるか?}

\begin{proof}
  By simultaneous induction on derivations.
 \qed
\end{proof}

\begin{lemma}[Stage Substitution]
	If $\G \V \mathcal{J}$, then $\G[\beta\mapsto B] \V \mathcal{J}[\beta\mapsto B]$.  Similarly, if $\G \V \mathcal{J}$, then $\V \G[\beta\mapsto B]$.
\end{lemma}

\begin{proof}
  By simultaneous induction on derivations.
  \qed
\end{proof}

% Inversion Lemma

The following Inversion Lemma is needed to prove main theorems.
As usual~\cite{TAPL}, Inversion Lemma enables us to infer the types of subterms of a term from the type of the term.

\begin{lemma}[Inversion]\ 
  \begin{enumerate}
  \item If $\G \V (\lambda x:\sigma.M) : (\Pi x:\sigma'.\tau)@A$,
    then $\G \V \sigma \E \sigma'@A$ and $\G ,x:\sigma@A\V M:\tau @A$.
  \item If $\G \V \TB_\alpha M : \TW_\alpha \tau @A$ then $\G \V M : \tau @A$.
  \item If $\G \V \Lambda\alpha.M : \forall\alpha.\tau @A$ then $\G \V M : \tau @A$ and $\alpha \notin \FTV(\G) \cup \FV(A)$.
  \end{enumerate}
\end{lemma}

\begin{proof}
  Each item is strengthen by statements about type equivalence.
  For example, the first statement is augmented by
  \begin{quotation}
    If $\G \V \rho \E (\Pi x:\sigma.\tau) : K @A$, then there exist
    $\sigma'$ and $\tau'$ such that $\rho = \Pi x:\sigma'.\tau'$ and
    $\G \V \sigma \E \sigma' : K @A$ and
    $\G, x:\sigma@A\V \tau \E \tau' : J @A$.
  \end{quotation}
  and its symmetric version.  Then, they are proved simultaneously by induction on derivations.

  Similarly for the others.
 \qed
\end{proof}

% Preservation

Thanks to Substitution Lemmas and Inversion Lemmas, we can prove Preservation easily.

\begin{theorem}[Preservation]
	If $\G\V M:\tau @A$ and $M \longrightarrow M'$, then $\G\V M':\tau @A$.
\end{theorem}

\begin{proof}
	First, there are three cases for $M \longrightarrow M'$.
	They are $M \longrightarrow_\beta M'$, $M \longrightarrow_\Lambda M'$, and $M \longrightarrow_\blacklozenge M'$.
	For each case, we can use straightforward induction on derivations. \AI{何の derivation?}
	Difficult cases are \TApp, \TTBL, and \TIns.
	We need Inversion Lemmas for them.
\end{proof}

% Strong Normalization

Strong Normalization is also an important property, which guarantees that
no typed term has an infinite reduction sequence.
We prove this thorem by translating \LMD to the symply typed lambda calculus~\cite{Tsukada} \AI{what is a standard citation for a proof of SN of a dependently typed calculus?}.

\begin{theorem}[Strong Normalization]
	If $\G\V M_1:\tau@A$ then there is no infinite sequence $(M_i)_{i\ge1}$ of terms such that
	$M_i \longrightarrow M_{i+1}$ for $i\ge 1$.
\end{theorem}

\begin{proof}
	In order to prove this theorem, we define a translation $(\cdot)^\natural$ from \LMD\ to the symply typed lambda calculus.
	Second, we prove the $\natural$-translation preserves typing and reduction.
	Then, we can prove Strong Normalization of \LMD from Strong Normalization of the symply typed lambda calculus.
\end{proof}

Confluence is a property that any reduction sequences from one typed term converge.
Since we have proved Strong Normalization, we can use Newman's Lemma~\cite{} to prove Confluence.

\begin{theorem}[Confluence]
	For any term $M$, if $M \longrightarrow^* M'$ and $M \longrightarrow^* M''$ then
	there exists $M'''$ that satisfies $M' \longrightarrow^* M'''$ and $M'' \longrightarrow^* M'''$.
\end{theorem}

\begin{proof}
  We can easily show Weak Church-Rosser.  Use Newman's Lemma.
	% Because we proved Strong Normalization of \LMD, 
	% we can use Newman's lemma to prove Confluence of \LMD.
	% Then, what we must show is Weak Church-Rosser Property now.
	% When we consider two different redexes in a \LMD term, they can only be disjoint, or one is a part of the other.
	% In short, they are never overlapped each other.
	% So, we can reduce one of them after we reduce another.
\end{proof}

Now, we turn our attention to proving Progress for staged semantics.
The following lemma Unique Decomposition ensures that, every typed
term is either a value or can be uniquely decomposed to an evaluation
context and a redex, ensuring that a well typed term is not
immediately stuck and the staged semantics is deterministic.

\begin{theorem}[Unique Decomposition]
  If $\G$ does not have any variable declared at stage $\varepsilon$ 
  and $\G \V M : \tau @ A$, then either
  \begin{enumerate}
  \item $ M \in V^A$, or
  \item there uniquely exist $B, E^A_B$, and $R^B$ such that $M = E^A_B[R^B]$.
  \end{enumerate}
\end{theorem}

\begin{proof}
  We can prove by straightforward induction on typing derivations.
  Main cases are \TApp, \TTBL, and \TIns.
  We need Inversion Lemmas for them.
\end{proof}

In a usual setting, the type environment $\G$ has to be empty, namely
the term has to be closed.  The condition is relaxed here because
variables at stages higher than \(\varepsilon\) are considered
symbols.  In fact, this relaxation is required for proof by induction
to work.

% This theorem is important because it guarantees
% that the evaluation context decides a redex to reduce deterministically.
% Specifically speaking, this theorem guarantee that 
% when you write a interpreter using the evaluation context of \LMD,
% your interpreter works just as intended.

Progress is a collorary of Unique Decomposition.

\begin{corollary}[Progress]
	If $\G$ does not have any variable declared at stage $\varepsilon$ and $\G \V M : \tau  @ A$ then
	$ M \in V^A $ or $M'$ exists such that $M \longrightarrow_s M'$.
\end{corollary}

\AI{Don't we prove $\longrightarrow_s$ is a subrelation of $\longrightarrow$?}