### 2019-06-22 text/conclusion.tex
Spellcheck (except proof.tex)

% !TEX root = ../main.tex

\section{Conclusion \label{sec:conclusion}}

We have proposed a new multi-stage calculus \LMD with dependent types.
\LMD enables finer-grained typing of multi-stage programming comparing to existing type systems.
This extension prevents some bugs relating to generated functions with multi-stage programming,
such as \verb|vadd| function for vectors of a fixed length.
We proved properties of \LMD including preservation, confluence, 
strong normalization for full reduction, and progress for staged reduction.
The technical difficulty of extension is how to deal with the symbol of CSP $\%$ in the equation rules.
We resolve this problem by ignoring $\%$ under some condition.

Developing a typechecking algorithmic for \LMD is left for future work.
They would be complicated but possible 
because we can realize them by defining a new weak head normal form which contains $\%$ erasing rule.