### 2019-07-30 reply-for-review.txt
Reflected comments from John.

Thank you for your detailed and valuable comments.

We will incorporate all feedback and comments, including those not addressed
here. We will now respond to some of the key comments and questsions from each
reviewer.

------------------ Reviewer 1 -------------------------

Comment on the last sentence of the abstract:
This sentence refers to the "vadd" example and handling of the symbol of CSP in
the type equivalence relation. We think they are an important part of this
paper. We'll try to improve but probably will not remove it.

Comment on the last sentence before Section 3.2:
This statement is a conjecture but we think it is not difficult to prove.
We will clarify this point.

Comment on design choice to make type-level CSP:
As we don't have staged semantics for types, explicit type-level CSP ultimately
complicates the system.

Comment on the proof of Strong Normalisation:
Yes, SN for a calculus with pattern-matching can be proved as suggested.  We
are not sure how the progress property contributes to the discussion here,
though: what we would like to prove is a simulation property, which assumes
one-step (successful) reduction in the source calculus and so the progress is
not used.  (Are we missing anything?)

------------------ Reviewer 2 -------------------------

Comment on "A metaprogramming framework for formal verification" (by Ebner et
        al.) and "Ur: Statically-Typed Metaprogramming with Type-Level Record
        Computation" (by Chlipala):
The first paper is loosely related in the sense that they are interested in
(automated) construction of well-typed terms but, in the MSP setting, the
presence of eval (which it does not care) often complicates the problem a lot.
We will include this discussion in the Related work section.
The second paper is also loosely related because it focuses on practical use
and does not contain a theory for metaprogramming.

Comment on "Program Logics for Homogeneous Generative Run-Time Meta-Programming"
    (by Berger and Tratt): 
Although this paper takes a different approach from ours, its main purpose is
similar to ours. The main difference is that their language is based on
lambda-Box, which can manipulate only closed code, whereas ours can deal
with open code.

------------------ Reviewer 3 -------------------------